---
title: "Preliminary Data Submission"
author: "Breh, Danna, Dahee, Jessi"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
```

## Preliminary Data Analysis Submission

# Load in Data
```{r}
pollinators <- read_csv("/Users/Breh/Desktop/ENT6702_Group_Project/Original_file_Breh/Pollinator_Obs_Data_24_ENT_6702.csv")
```
The data used is from Breh's thesis research. It is pollinator observation data collected from Summer 2024.

# Inspect the first few rows of the data
```{r}
head(pollinators)
```

# Ensure variables are assigned their correct data types
```{r}
pollinators <- pollinators %>%
  mutate(
    Location = as.factor(Location),              
    Application = as.factor(Application),        
    Date = as.Date(Date, format = "%Y-%m-%d"),   
    Treatment = as.factor(Treatment),            
    Zone = as.factor(Zone),                      
    Sample = as.factor(Sample),                
    Pollinator_category = as.factor(Pollinator_category),  
    Duration_of_visit = as.numeric(Duration_of_visit),
    Pollination_event = as.factor(Pollination_event)   
  )
```

# Subset the data to only include POST treatment data
```{r}
post_data <- pollinators %>%
  filter(Application == "Post") %>% 
  droplevels() 
```

# Inspect the first few rows of the subset data
```{r}
str(post_data)
summary(post_data)
head(post_data)
tail(post_data)
view(post_data)
```

# Viewing data and dealing with NA values 
```{r}
post_data %>% group_by(Pollinator_category) %>% summarise (Means= mean(Duration_of_visit), SD = sd(Duration_of_visit), max_sep = max(Duration_of_visit)) %>% arrange(Means)
mean(post_data$"Duration_of_visit")
```

# Boxplot of a specific variable
```{r}
ggplot(post_data, aes(x = Pollinator_category, y = Duration_of_visit)) + 
  geom_boxplot() + theme_minimal()

ggplot(post_data, aes(x = Treatment, y = Duration_of_visit)) + 
  geom_boxplot() +
  theme_minimal()
```

# Boxplot with points 
```{r}
ggplot(post_data, aes(x = Pollinator_category, y = Duration_of_visit)) + 
    geom_boxplot() +  geom_jitter(width = 0.2, size = 2, color = "blue") +
    theme_minimal()

ggplot(post_data, aes(x = Treatment, y = Duration_of_visit)) + 
  geom_boxplot() + geom_jitter(width = 0.2, size = 2, color = "blue") +
  theme_minimal()
```

# Q-Q to see normality 
```{r}
qqnorm(post_data$Duration_of_visit)
qqline(post_data$Duration_of_visit, col = "red")
```

# Checking normality with histogram
```{r}
hist(post_data$Duration_of_visit)
```

# Checking normality with shapiro test
```{r}
shapiro.test(post_data$Duration_of_visit)
```

# Checking homogeneity of variance
```{r}
var_check <- post_data %>% group_by(Pollinator_category) %>% summarise(Variance = var(Duration_of_visit))
var_check$Variance[2]/var_check$Variance[1]
```

# Conclusion: 
The data does not have a normal distribution of its numerical variable duration of visit. 

# Transforming data with log transformation to see impact on normality
```{r}
normalized_post_data<-log(post_data$Duration_of_visit)
qqnorm(normalized_post_data)
qqline(normalized_post_data,col = "red")
shapiro.test(normalized_post_data)
```

# Final conclusion: 
Data closer to normal, however, still not fully.